Refunctionalization of Abstract Abstract Machines (Functional Pearl)

### 2018-06-17 paper/paper.tex
reformat code
fix cache in code and paper

%------------------------------------------------------------------------------------------------------------
\begin{lstlisting}
                                    ae $\in$ AExp ::= ∂x | lam
                                     x $\in$ Var   (variable names)
                                   lam $\in$ Lam ∂::= (lambda (x) e)
                                     e $\in$ Exp ∂::= ae | (let ([x (ae ae)]) e)
\end{lstlisting}

@@

\begin{lstlisting}
       val mtCache = Cache(Store[Config, VS](Map()), Store[Config, VS](Map()))
       def analyze(e: Expr) = {
         def iter(cache: Cache): Ans = {
           val Ans(vss, new_cache) = aeval(e, mtEnv, mtStore, mtTime, cache, { case Ans(vss, cache) => 
               Ans(vss, cache.outUpdate(Config(e, mtEnv, mtStore, mtTime), vss))
           })
           if (new_cache.out == cache.out) { Ans(vss, new_cache) }
           else { iter(Cache(new_cache.out, new_cache.out)) }
         }
         iter(mtCache)
       }
\end{lstlisting}

Finally, the @analyze@ function (as the entrance of the analysis) does a looping iteration
to find the fixed-point on caches starting from an empty cache.
If the @out@ cache of this iteration is equivalent to the @out@ cache from the last iteration,
then we have reached a fixed-point that over-approximates the concrete evaluations,
and thus can be returned.
Otherwise, the next iteration with the new @out@ cache will be activated.
Note that the continuation of the initial call to @aeval@ sets up the values for the
initial configuration in the final cache.

@@

We have multiple choices of how to proceed:
\begin{itemize}
  \item By monadifying the continuations, we obtain the abstracting
    definitional interpreters in monadic style which are what \citeauthor{darais2017abstracting}
    describe \cite{darais2017abstracting}.
  \item By representing the extended continuation-passing style with delimited controls,
    we derive a new form of abstract interpreters in direct-style.
  \item In fact, we may also just use the same caching algorithm but with side effects such as
    assignments and mutations to update the cache, and then achieve the same definitional 
    interpreter with pushdown control-flows. In this case, nondeterminism can be easily 
    handled via @for@ comprehension.
\end{itemize}

@@

\begin{lstlisting}
             def analyze(e: Expr) = {
               def iter(cache: Cache): Ans = {
                 val Ans(vss, anscache) = reset { aeval(e, mtEnv, mtStore, mtTime, cache) }
                 val new_cache = anscache.outUpdate(Config(e, mtEnv, mtStore, mtTime), vss)
                 if (new_cache.out == cache.out) { Ans(vss, new_cache) }
                 else { iter(Cache(new_cache.out, new_cache.out)) }
               }
               iter(mtCache)
             }
\end{lstlisting}

>>>>

%------------------------------------------------------------------------------------------------------------
\begin{lstlisting}
                                     x $\in$ Variables
                                    ae $\in$ AExp ::= ∂x | lam
                                   lam $\in$ Lam ∂::= (lambda (x) e)
                                     e $\in$ Exp ∂::= ae | (let ([x (ae ae)]) e)
\end{lstlisting}

@@

\begin{lstlisting}
val mtCache = Cache(Store[Config, VS](Map()), Store[Config, VS](Map()))
def analyze(e: Expr) = {
  def iter(cache: Cache): Ans = {
    val Ans(vss, new_cache) = aeval(e, mtEnv, mtStore, mtTime, cache, { case Ans(vss, cache) => 
        Ans(vss, cache.outUpdate(Config(e, mtEnv, mtStore, mtTime), vss))
    })
    if (new_cache.out == new_cache.in) Ans(vss, new_cache)
    else iter(Cache(new_cache.out, Store[Config, VS](Map())))
  }
  iter(mtCache)
}
\end{lstlisting}

% TODO: explain more on how caches work
Finally, the @analyze@ function (as the entrance of the analysis) does a looping iteration
to find the fixed-point on caches starting from an empty cache.
If the @out@ cache of this iteration is equivalent to its @in@ cache, which means no more 
fact was discovered during this iteration, then we have reached a fixed-point that 
over-approximates the concrete evaluations, and thus can be returned.
Otherwise, the next iteration with the new @out@ cache installed at the @in@ cache position 
will be activated; in the meanwhile, an empty cache will be used as the @out@ cache.
Note that the continuation of the initial call to @aeval@ sets up the values for the
initial configuration in the final cache.

@@

We have multiple choices of how to proceed:
\begin{itemize}
  \item By monadifying the continuations, we obtain the abstracting
    definitional interpreters in monadic style which are what \citeauthor{darais2017abstracting}
    describe \cite{darais2017abstracting}.
  \item By representing the extended continuation-passing style with delimited control operators
    such as @shift@ and @reset@, we derive a new form of abstract interpreters in direct-style.
  \item In fact, we may also just use the same caching algorithm but with side effects such as
    assignments and mutations to update the cache, and then achieve the same definitional 
    interpreter with pushdown control-flows. \citeauthor{friedman-ai} in a tutorial on abstract 
    interpreters use this style to model caches and updates \cite{friedman-ai}.
    In this case, nondeterminism can be easily handled via @for@ comprehension. 
\end{itemize}

@@

\begin{lstlisting}
def analyze(e: Expr) = {
  def iter(cache: Cache): Ans = {
    val Ans(vss, anscache) = reset { aeval(e, mtEnv, mtStore, mtTime, cache) }
    val new_cache = anscache.outUpdate(Config(e, mtEnv, mtStore, mtTime), vss)
    if (new_cache.out == new_cache.in) Ans(vss, new_cache)
    else iter(Cache(new_cache.out, Store[Config, VS](Map())))
  }
  iter(mtCache)
}
\end{lstlisting}
